A Particular Appetite: Cosmological Hydrodynamic Simulations of Preferential Accretion in the Supermassive Black Holes of Milky Way Size Galaxies

### 2016-05-08 ms.tex
Jillian Edits: Done; Expanding Results for HR h258

We used Charm N-body GrAvity solver, ChaNGa, a high resolution version of the same volume. The results of this paper, both from Gasoline and ChaNGa, utilize the same initial conditions but the latter is simulated at a higher resolution allowing the following new physics to be included: metal line cooling \citep{Christensen2012,Shen2010}, gas-phase and dust grain H_2 formation \citep{Christensen2014}, star formation dependence on H$_2$, and H$_2$ destruction from Lyman-Werner radiation. Other changes include a spline force softening lengths of 174 pc and initial gas particle masses of $2.7 \times 10^4 M_{\odot}$. Star particles are created with 30\% of their parent gas particle mass; therefore, a maximum initial mass of 8100 $M_{\odot}$. Each galaxy contains about 5 million DM particles inside the virial radius at z=0 and over 14 million DM, star, and gas particles total. The resolution of both force and mass in these simulations is comparable to the ``Eris'' simulation which has one of the highest resolutions for an N-body+SPH cosmological simulation of a Milky Way-mass galaxy so far produced.  At z=9, a uniform UV background is applied to simulate the cosmic reionization energy in a variation of \cite{Haardt2012}.

@@

The Amiga Halo Finder identifies all of the halos in the simulations by identifying the virial radii of the halos based on an overdensity using the criterion for a flat universe. \citep{Knebe2001,Knollmann2009,} We select the primary halo by determining the most massive galaxy at z=0 and locating the central SMBH. For our Gasoline simulations, the primary halo in h258 had a final mass of [get mass],

EDIT: Include specs: mass, formation z, N (for primary halos? Ask Kelly what she means.)

@@

\subsection{Low Resolution h258: Resolution Test}

The galaxy, h258, is characterized by an active merger history including a major merger at z=1. The merger can be seen affecting the cumulative black hole mass (Figure \ref{h258allmassgas}) as well as the gas fractions within the both the SMBH and galaxy (Figure \ref{h258numfrac}). 
Though Figure \ref{h258allmassgas} shows that the primary source of mass gain at high redshifts is through mergers with other black holes, there is a significant increase in the amount of gas adding to the mass of the SMBH starting at z $\sim$ 1.

Similarly, Figure \ref{h258numfrac} shows that unshocked gas (blue) makes up the majority of gas entering the galaxy (solid lines) at early times; nevertheless, around the time of the merger (z $\sim$ 1) a marked change occurs in the gas fractions within the black hole (dashed lines) where merger gas (green) becomes the main component of total gas.

Figure \ref{h258piefrac} depicts the fractions of total gas in the galaxy (a) and the SMBH (b) at z=0, again differentiated by the gas particles' origins. The galaxy has a mass nearly half comprised of gas entering the galaxy through unshocked, smooth accretion (48 \%), with 39 \% of the gas entering through mergers. The smallest fractions of the total gas are comprised of shocked gas (9 \%) and ``early''gas which exiisted at formation makes up the final, smallest fraction (4 \%). Unlike the galaxy, nearly 65 \% of the gas accreted by the central SMBH was gas accreted via mergers, while only a quarter (26 \%) is comprised of unshocked, smoothly accreted gas. Shocked gas makes up the last 9 \% of total gas entering the SMBH, along with a negligible (less than 1\%) amount of ``early'' gas. This result shows explicitly that the SMBH more readily accretes gas gained through mergers which is \textit{contrary} to the previous results of \cite{Bellovary2013}.

We examine the angular momentum of the gas at the moment it enters the galaxy for clues to the understand the ready accretion of merger accreted gas. (Figure \ref{h258angmom}) A cumulative distribution of the angular momentum of entering gas, differentiated based on method of entrance into the galaxy, shows that gas entering the SMBH (dashed lines) has an overall lower angular momentum than gas entering galaxy (solid lines). This is true whether the gas enters via unshocked smooth accretion, becomes shocked upon entry, or enters through a galaxy merger (blue, red, green, respectively).

 By examining the angular momentum of the gas accreted onto the SMBH at the moment it enters the halo, we were able to determine that a significant preference exists for gas with lower angular momentum to enter the SMBH. In contrast, \cite{Bellovary2013} found a \textit{slight} preference for gas of a lower angular momentum to enter the SMBH. From our results we determine that the gas attained through mergers tends to enter the halo with a lower angular momentum which may explain the SMBH's preference for merger gas.

\subsection{Low Resolution h277: Resolution Test}

Our quiescent galaxy, h277, was characterized by few mergers in its past, its final major merger occuring around z $\sim$ 3. This merger is visible in Figure \ref{h277allmassgas}, though more subtle than the influx of gas in h258. In h277, gas seemed to be gained at a more steady rate and even at fairly high redshifts comprises a larger portion of the black holes total mass than seen in the previous galaxy. The gas fractions (Figure \ref{h277numfrac}) in the galaxy (solid lines) and SMBH (dashed lines) are also markedly different than that of a galaxy with an active merger history. While it appears neither merger-gained or unshocked gas (green and blue, respectively) particularly dominates at higher redshifts in the galaxy, merger gas becomes prominent in the SMBH at times when mergers occur. Again, this implies some physical process allowing merger gas to readily fuel black hole growth. Though, it is important to note that by z $\sim$ 0, smoothly accreted, unshocked gas dominates in both the SMBH and host galaxy.

This result can be more clearly seen in Figure \ref{h277piefrac}. Similar to Bellovary et. al. 2013 study, which found SMBH accretion in high mass, high redshift galaxies mirror their hosts in terms of accretion, h277's SMBH accretes a similar fraction of each type of gas, comparable to its host galaxy. Smoothly accreted, unshocked gas represents about half of the total gas in both the SMBH and the galaxy. Merger-gained gas also maintains about 30\% of both. There is an increase in the shocked gas accreted by the SMBH; however, we argue this may be due to supernova contamination in the simulation. So by present day, h277, unlike the active h258, found no strong preference for the SMBH to prefer any one kind of gas in its accretion. Like the SMBH in the study of \cite{Bellovary2013}, the same fraction of gas entering the halo would then be accreted onto the SMBH.

As in the previous case, we examined the angular momentum of all the gas particles accreted by the SMBH and host galaxy at the moment of entry. (Figure \ref{h277angmom}) It is immediately apparent that the merger gas accreted by both the black hole and its host has a lower angular momentum overall than the smoothly accreted gas; however, unlike we found in h258, the h277 galaxy accretes gas with lower angular momentum than that of its SMBH.

As h277 and h258 are similar in size and appearance at z=0, these differences indicate that the varying histories of these simulations plays a role in the evolution of these types of galaxies and their central SMBHs. We also note that mergers between galaxies appear to be important events when considering the gas accretion of SMBHs.

\subsection{High Resolution h258}

[Still need to write; needs results first]

(Assuming high rez finds same preferential accretion?)\\
$\bullet$ Just as we found in the lower resolution simulation, the high resolution sees a distinct preference for accreting gas that has been gained through the many mergers in its history. \\
$\bullet$ (MAYBE. We don't know if this is true yet.)\\

@@

The final fraction of ``early'' gas in the galaxy at z=0 is less than four percent for both h258 and h277, a negligible amount.\\

@@

>>>>

We used Charm N-body GrAvity solver, ChaNGa, a high resolution version of the same volume. The results of this paper, both from Gasoline and ChaNGa, utilize the same initial conditions but the latter is simulated at a higher resolution allowing the following new physics to be included: metal line cooling \citep{Christensen2012,Shen2010}, gas-phase and dust grain H$_2$ formation \citep{Christensen2014}, star formation dependence on H$_2$, and H$_2$ destruction from Lyman-Werner radiation. Other changes include a spline force softening lengths of 174 pc and initial gas particle masses of $2.7 \times 10^4 M_{\odot}$. Star particles are created with 30\% of their parent gas particle mass; therefore, a maximum initial mass of 8100 $M_{\odot}$. Each galaxy contains about 5 million DM particles inside the virial radius at z=0 and over 14 million DM, star, and gas particles total. The resolution of both force and mass in these simulations is comparable to the ``Eris'' simulation which has one of the highest resolutions for an N-body+SPH cosmological simulation of a Milky Way-mass galaxy so far produced.  At z=9, a uniform UV background is applied to simulate the cosmic reionization energy in a variation of \cite{Haardt2012}.

@@

The Amiga Halo Finder identifies all of the halos in the simulations by identifying the virial radii of the halos based on an overdensity using the criterion for a flat universe. \citep{Knebe2001,Knollmann2009,} We select the primary halo by determining the most massive galaxy at z=0 and locating the central SMBH. For our Gasoline simulations, both the primary halo in h258 and h277 had final masses on the order of $10^7 M_{\odot}$ and a formation redshift of z~4.

EDIT: Include specs: mass, formation z, N (just for primary halos? Ask Kelly what she means.)

@@

\subsection{Gasoline - Low Resolution h258: Resolution Test}

The galaxy, h258, is characterized by an active merger history including a major merger at z=1. The merger can be seen affecting the cumulative black hole mass (Figure \ref{h258allmassgas}) as well as the gas fractions within the both the SMBH and galaxy (Figure \ref{h258numfrac}). 
Figure \ref{h258allmassgas} shows h258's central SMBH's cumulative mass as a function of time (lower axis) and redshift (upper axis). The black dashed line indicates the total cumulative BH mass (including both mass from gas and BH mergers), while the black solid line indicates only the total gas mass accreted by the BH. The blue dot-dashed line indicates the gas mass accreted via unshocked gas. The green solid line indicates the gas mass accreted through mergers. The red dashed line indicates gas mass that was shocked upon entry into the halo. Though this figure shows that the primary source of mass gain at high redshifts is through mergers with other black holes, there is a significant increase in the amount of gas adding to the mass of the SMBH starting at z $\sim$ 1, where we know a major merger takes place. \textbf{[Reference merger tree here once you've added it.]} Similarly, Figure \ref{h258numfrac} shows that unshocked gas (blue) makes up the majority of gas entering the galaxy (solid lines) at early times; nevertheless, around the time of the merger (z $\sim$ 1) a marked change occurs in the gas fractions within the black hole (dashed lines) where clumpy gas (green) becomes the main component of total gas.

Figure \ref{h258piefrac} depicts the fractions of total gas in the galaxy (a) and the SMBH (b) at z=0, again differentiated by the gas particles' origins. Blue, green, and red distinguish cold, clumpy, and shocked, respectively. Yellow indicates gas that existed within the main halo upon formation. The galaxy has a mass nearly half comprised of gas entering the galaxy through unshocked, smooth accretion (48 \%), with 39 \% of the gas entering through mergers. The smallest fractions of the total gas are comprised of shocked gas (9 \%) and ``early''gas which existed at formation makes up the final, smallest fraction (4 \%). Unlike the galaxy, nearly 65 \% of the gas accreted by the central SMBH was gas accreted via mergers, while only a quarter (26 \%) is comprised of unshocked, smoothly accreted gas. Shocked gas makes up the last 9 \% of total gas entering the SMBH, along with a negligible ($<$ 1\%) amount of ``early'' gas. \textbf{So Figure \ref{h258piefrac} shows explicitly that the SMBH more readily accretes gas gained through mergers.} This result is \textit{contrary} to the previous results of \cite{Bellovary2013}! This result is surprising, but important, as it allows us to consider that a different physical mechanism characterizes the gas entering this MW-type galaxy than those affecting the gas accretion of the high mass galaxies examined by \cite{Bellovary2013}.

We examine the angular momentum of the gas at the moment it enters the galaxy for clues to the understand the ready accretion of merger accreted gas. (Figure \ref{h258angmom}) A cumulative distribution of the angular momentum of entering gas, differentiated based on method of entrance into the galaxy, shows that gas entering the SMBH (dashed lines) has an overall lower angular momentum than gas entering galaxy (solid lines). This statement is true whether the gas enters via unshocked smooth accretion, becomes shocked upon entry, or enters through a galaxy merger (blue, red, green, respectively). Therefore, we can infer from this plot that lower angular momentum gas more readily flows into the central region of the host galaxy where it can be accreted by the black hole. In contrast, \cite{Bellovary2013} found a \textit{slight} preference for gas of a lower angular momentum to enter the SMBH. The question still remains as to why clumpy gas in particular appears most prominently in the SMBH at z = 0. The answer to this conundrum may lie in the major merger that characterizes this galaxy, so we look specifically at the gas accreted during the merger.

Figure \ref{h258angmom_merger} explicitly shows that at the time of the merger, clumpy gas (green) has the lowest angular momentum entering the galaxy (solid lines) or the SMBH (dashed lines). Both cold gas and shocked gas (blue and red, respectively) are entering the galaxy and SMBH with significantly higher angular momentum. Therefore, as we've already seen that the SMBH tends to preferentially accrete gas with lower angular momentum, we determine that the significantly lower angular momentum with which the clumpy gas enters the galaxy at this timestep, allows for nearly all of the clumpy gas to then funnel directly into the SMBH. It is clear through this result the physical properties afecting the galaxy through this major merger play an important role in allowing for the direct fueling of the SMBH at its center.

\subsection{Gasoline - Low Resolution h277: Resolution Test}

Our quiescent galaxy, h277, was characterized by few mergers in its past, its final major merger occuring around z $\sim$ 3. This merger is visible in Figure \ref{h277allmassgas}, though more subtle than the influx of gas in h258. (Linestyles as in Figure \ref{h258allmassgas}.) In h277, gas seemed to be gained at a more steady rate and at fairly high redshifts comprises a larger portion of the BH's total mass than seen in the previous, more active galaxy. The gas fractions (Figure \ref{h277numfrac}) in h277 (solid lines) and SMBH (dashed lines) are also markedly different than that of a galaxy with an active merger history. While it appears neither clumpy or unshocked gas (green and blue, respectively) particularly dominates at higher redshifts in the galaxy, merger gas becomes prominent in the SMBH at times when mergers occur. \textbf{[Ref merger tree when included.]} Again, this implies some physical process allowing merger gas to readily fuel black hole growth. Though, it is important to note that by z $\sim$ 0, smoothly accreted, unshocked gas again dominates in both the SMBH and host galaxy. As in our more active galaxy, it appears that the role of clumpy gas as it fuels the SMBH becomes important when mergers affect the galaxy; however, with h277 having a significantly less active merger history, unshocked, smooth accretion inevitably dominates the total gas supply.

This result can be more clearly seen in Figure \ref{h277piefrac}. (Colors as in \ref{h258piefrac}.) Similar to Bellovary et. al. 2013 study, which found SMBH accretion in high mass, high redshift galaxies mirror their hosts in terms of accreted gas, h277's SMBH accretes a similar fraction of each type of gas, comparable to its host galaxy. Smoothly accreted, unshocked gas represents about half of the total gas in both the SMBH and the galaxy. Clumpy gas also maintains about 30\% of both, while an increase in shocked gas occurs, from 13 \% in the galaxy to 22 \% in the SMBH. %; however, we argue this may be due to supernova contamination in the simulation. 
So by present day, h277 found no strong preference for the SMBH to prefer any one kind of gas in its accretion. Like the SMBHs in the study of \cite{Bellovary2013}, nearly the same fractions of gas entering the halo would then be accreted onto the SMBH (with the exception of shocked gas). Therefore, this may imply that without some physical driver, like the major merger of h258, clumpy gas is unlikely to dominate the gas accreted into the SMBH.

As in the previous case, we examined the angular momentum of all the gas particles accreted by the SMBH and host galaxy at the moment of entry. (Figure \ref{h277angmom}; Linestyles as in \ref{h258angmom}) Like before, it is immediately apparent that the merger gas accreted by both the black hole and its host has a lower angular momentum overall than the smoothly accreted gas; however, unlike we found in h258, the h277 galaxy accretes gas with lower angular momentum than that of its SMBH. This means that even though some of the gas in the host halo had lower angular momentum, that gas didn't find its way into the central region to accrete onto the BH. This may again imply that some physical phenomena is required to drive those gases towards the center.

As h277 and h258 are similar in size and appearance at z=0, these differences indicate that the varying histories of these simulations appears to play an important role in the evolution of these types of galaxies and their central SMBHs. We also note that while mergers may not be the only physical mechanisms by which gas can be funneled into the centers of galaxies, mergers between galaxies clearly play an important role when considering the gas accretion of SMBHs.

\subsection{ChaNGa - High Resolution h258}

Just as we found in the lower resolution simulation, the high resolution sees a distinct preference for accreting gas that has been gained through the many mergers in its history.

@@

@@

\begin{figure}
\centerline{\resizebox{0.75\hsize}{!}{\includegraphics[angle=0]{h258_am_tsnew_228}}}
\caption[]{ Cumulative distribution of angular momentum of the gas particles accreted onto h258 at the time of the major merger (z ~ 1). There are about 95,000 gas particles accreted at this timestep, which gas particles accreted onto the main halo and central black hole distinguished by solid and dashed lines. The green, blue, and red lines indicate clumpy, unshocked, and shocked gas, respectively.}
\label{h258angmom_merger} 
\end{figure}
